Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-01 ms.tex
brak4

If heating in the nuclei of elliptical galaxies is similar to that produced by the average star formation history of galaxies with similar mass, then to first order we predict that $\langle L_{X}/L_{\rm edd}\rangle$ should increase with black hole mass, though only weakly for $M_{\bullet} \lesssim 3\times 10^{8}M_{\odot}$.  This result is in tension with the observed down-sizing trend that the average Eddington ratio is observed to decrease (albeit also weakly) with $M_{\bullet}$.  However, in making such comparisons it is important to keep in mind the enormous uncertainty in calculating the luminosity of the accretion flow close to the black hole in a single waveband based on the accretion rate on larger scales.  Black hole mass-dependent efficiency factors could enter that we are not taking into account.  For instance, if disk outflows carry away a substantial fraction of the infalling mass before it reaches sufficiently small radii to produce the observed X-rays (e.g.~\citealt{Blandford&Begelman99}), then the angular momentum of the infalling gas (and hence that of stars in the galactic nucleus) might also affect the X-ray luminosity, potentially in a way that could depend systematically on the black hole mass.

Perhaps a more readily addressable test is whether a steady-state accretion picture can explain the large large scatter of typically 2-3 orders of magnitude observed in $L_{X}/L_{\rm edd}$.  Scatter can result from at least two effects (1) the order of magnitude difference between $\dot{M}$ for core and cusp galaxies at fixed $v_w$ and $M_{\bullet}$ (Fig.~\ref{fig:bh_growth}) will result in a similar scatter of $\dot{M}$ due to the variation in the Nuker stellar density profile $\Gamma$ at fixed $M_{\bullet}$; (2) the strong sensitivity of the accretion rate to the stellar wind velocity, $\dot{M}/\dot{M}_{\rm edd} \propto v_{w}^{-3.8(-2.4)}$ for core(cusp) galaxies, respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]) in combination with the significant dependence of $v_w$ on stochastic intermittency in the star formation history (Fig.~NICK).  The combination of (1) and (2) could well explain much of the scatter observed.

However, it is also possible that many of the galactic nuclei from the \citet{Miller+15} sample are in a thermally stable state.  The efficiency show in the right panel of Figure \ref{fig:miller} in some ways represents the maximum efficiency, since it does not take into account outflows from the accretion flow that would suppress the accreiton rate reaching large radii.  Nevertheless, many of the observed galaxies at low black hole masses of $M_{\bullet} \lesssim 10^{7}M_{\odot}$ lie well below the predicted average value for stable accretion (black line) and even approaching or above the maximally stable accretion rate $\dot{M}_{\rm TI}$ (yellow line) .  This suggests that these galaxies experience lower levels of heating than average galaxies of their mass, or that they are not accreting stably at all ($\S\ref{sec:stable}$).

Also clear from our results are that black holes accreting in AGN phase cannot be accreting in a steady

\subsection{Steady Accretion versus Cyclic Outbursts}

>>>>

If heating in the nuclei of elliptical galaxies is similar to that produced by the average star formation history of galaxies with similar mass, then to first order we predict that $\langle L_{X}/L_{\rm edd}\rangle$ should increase with black hole mass, though only weakly for $M_{\bullet} \lesssim 3\times 10^{8}M_{\odot}$.  This result is in tension with the observed down-sizing trend that the average Eddington ratio is observed to decrease (albeit also weakly) with $M_{\bullet}$.  However, in making such comparisons it is important to keep in mind the enormous uncertainty in calculating the luminosity of the accretion flow close to the black hole in a single waveband based on the accretion rate on larger scales.  Black hole mass-dependent efficiency factors could enter that we are not taking into account.  For instance, if disk outflows carry away a substantial fraction of the infalling mass before it reaches sufficiently small radii to produce the observed X-rays (e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}), then the angular momentum of the infalling gas (and hence that of stars in the galactic nucleus) might also influence the X-ray luminosity, in a way that could depend systematically on $M_{\bullet}$.

A more readily addressable test is whether a steady-state accretion picture is capable of explaining the large scatter of typically 2-3 orders of magnitude in the observed distribution of $L_{X}/L_{\rm edd}$ at fixed $M_{\bullet}$.  Scatter is expected to result from at least two effects: (1) the order of magnitude difference between $\dot{M}$ for core and cusp galaxies at fixed $v_w$ and $M_{\bullet}$ (Fig.~\ref{fig:bh_growth}) leads to a similar scatter of $\dot{M}$ due to the variation in the Nuker stellar density profile $\Gamma$ at fixed $M_{\bullet}$; (2) the strong sensitivity of the accretion rate to the stellar wind velocity, $\dot{M}/\dot{M}_{\rm edd} \propto v_{w}^{-3.8(-2.4)}$ for core(cusp) galaxies, respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]) in combination with the significant dependence of $v_w$ on stochastic intermittency in the star formation history (Fig.~NICK).  The combination of (1) and (2) is capable of explaining much of the observed sample.

It is also possible that many of the nuclei from the \citet{Miller+15} sample are not accreting in a steady fashion.  The ADAF efficiency assumed in the right panel of Figure \ref{fig:miller} in some ways represents the maximum value, since it does not take into account outflows from the accretion flow that would suppress the accreiton rate reaching large radii.  Nevertheless, many lower black hole masses of $M_{\bullet} \lesssim 10^{8}M_{\odot}$ show X-ray luminosities lie well above the predicted average value for stable accretion (black line), suggesting that these galaxies either experience lower levels of heating than average galaxies of their mass, or that they are not accreting stably at all ($\S\ref{sec:cycle}$).  Indeed, for the lowest mass holes, $M_{\bullet} \lesssim 10^{8}M_{\odot}$, $L_{\rm X}$ exceeds the one corresponding to the maximum thermally-stable accretion rate, $\dot{M}_{\rm TI}$ (yellow line).  Many of the upper limits on $L_{\rm X}$ at low $M_{\bullet}$ may in fact correspond to a separate ``thermally stable".  This possible dichotomy calls into question the use of simple extrapolations of power-law fits to the $L_{\rm X}(M_{\bullet}$) relationship to low $M_{\bullet}$ to constrain the occupation fraction of black holes in low mass galaxies (\citealt{Miller+15}).

\subsection{Steady Accretion versus Cyclic Outbursts}
\label{sec:cycle}
